---
title: "Effiecncy of relic DNA quantifaction"
author: "Jay T. Lennon, Mario E. Muscarella"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

Comparing efficiency of relic DNA "removal" using DNase and PMA

# Setup Work Environment
```{r, results='hide', warning=FALSE, message=FALSE}
#rm(list=ls())
getwd()
setwd("~/GitHub/relicDNA/code")
require("plyr")
require("grid")
require("png")
require("car")
require("bbmle")
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```

# Load data and calcualte corrected copy number
```{r}
eDNA.raw <- read.table("../data/eDNA_efficiency.txt", sep = "\t", header = T)


# Make new dataframe with corrected copy numbers
eDNA.corr <- data.frame(eDNA.raw, copies.corr)
```

# Take mean of technical replicates and sort
```{r}
# Use `aggregate` to return mean or triplicate qPCR runs

eDNA <- aggregate(eDNA.raw$copies ~ eDNA.raw$sample + eDNA.raw$treatment, eDNA.raw, mean)

# Sort by sample number
eDNA <- eDNA[order(eDNA[,1]) ,]

# Rename columns
colnames(eDNA) <- c("sample", "treat", "copies")
```

# Number of copies added by spike
```{r}
spike <- eDNA[ which(eDNA$treat == 'spike'),]
control <-  eDNA[ which(eDNA$treat == 'control'),]
spiked.DNA <- spike$copies - control$copies
```

# PMA efficiency
```{r}
# amount of DNA remaining after PMA treamtnet
PMA <- eDNA[ which(eDNA$treat == 'PMA'),]
PMA.remain <- PMA$copies/spiked.DNA
PMA.efficiency <- 1-PMA.remain
PMA.mean <- mean(PMA.efficiency)
PMA.sem <-sem(PMA.efficiency)
PMA.cv <- sd(PMA.efficiency)/mean(PMA.efficiency) # CV

# Mean PMA efficiency is 70% but there's a low outlier of .36
```

# DNase efficiency
```{r}
# amount of DNA remaining after DNase treamtnet
DNase <- eDNA[ which(eDNA$treat == 'DNase'),]
DNase.remain <- DNase$copies/spiked.DNA
DNase.efficiency <- 1-DNase.remain
DNase.mean <- mean(DNase.efficiency)
DNase.sem <-sem(DNase.efficiency)
DNase.cv <- sd(DNase.efficiency)/mean(DNase.efficiency)

# Mean DNase efficiency is 98% 
```

# Does efficiency differ between methods?
```{r}
treat.PMA.id <- c("PMA", "PMA", "PMA", "PMA")
treat.PMA <- data.frame(PMA.efficiency, treat.PMA.id)
as.data.frame(treat.PMA)
colnames(treat.PMA) <- c("efficiency", "treat")

treat.DNase.id <- c("DNase", "DNase", "DNase", "DNase")
treat.DNase <- data.frame(DNase.efficiency, treat.DNase.id)
as.data.frame(treat.DNase)
colnames(treat.DNase) <- c("efficiency", "treat")

t.test.data <- rbind(treat.PMA, treat.DNase)

efficiency.test <- t.test(t.test.data$efficiency ~ t.test.data$treat)
summary(efficiency.test)

# conclusion is that PMA and DNase are equally and highly efficient
```


# Plot efficiencies
```{r}
# Make table with summary stats
PMA.sum <- data.frame(PMA.mean, PMA.sem)
colnames(PMA.sum) <-c("mean", "sem")
DNase.sum <- data.frame(DNase.mean, DNase.sem)
colnames(DNase.sum) <-c("mean", "sem")
eDNA.sum <- rbind(PMA.sum, DNase.sum)

# Make bar plot with error bars by environment
```{r, eval=F}
png(filename="../figures/efficiency.png",
    width = 800, height = 800, res = 96*2)

bp <- barplot(eDNA.sum$mean, ylim =c(0, 1.1),
              pch = 15, cex = 1.25, las = 1, cex.lab = 1.25, cex.axis = 1,
              col = "gray90", axis.lty = 1, lwd = 2, xlab = NA,
              ylab = "Efficiency of relic DNA removal (%)",
              names.arg = c("PMA", "DNase"), cex.names = 0.9)
              box(lwd = 2)
arrows(x0 = bp, y0 = eDNA.sum$mean, y1 = eDNA.sum$mean - eDNA.sum$sem,
       angle = 90,  length = 0.1, lwd = 2)
arrows(x0 = bp, y0 = eDNA.sum$mean, y1 = eDNA.sum$mean + eDNA.sum$sem,
       angle = 90, length=0.1, lwd = 2)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/efficiency.png")
grid.raster(img)
```
