---
title: "eDNA"
author: "Jay T. Lennon, Mario E. Muscarella, ..."
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

A model approach to relic DNA.

# Initial Setup
```{r}
#rm(list=ls())
setwd("~/GitHub/relicDNA/code")
library("vegan")
library("png")
library("grid")
set.seed(Sys.time())
```

# Sampling Mixing Models
## Different Evenness
```{r}
gamma1 <- rlnorm(n=10000, meanlog = 1, sdlog = 0.98)
gamma1 <- gamma1[rev(order(gamma1))]
gamma2 <- rlnorm(n=10000, meanlog = 1, sdlog = 2)
gamma2 <- gamma2[rev(order(gamma2))]
gamma3 <- rlnorm(n=10000, meanlog = 1, sdlog = 0.1)
gamma3 <- gamma3[rev(order(gamma3))]

# Simple Plot
layout(1)
plot(log10(gamma1), col = "forestgreen",
     main = "Species Abundance Distributions",
     xlab = "Species Index", ylab = "Log Probability",
     las = 1)
points(log10(gamma2), col = "red")
points(log10(gamma3), col = "blue")

# Define OTUs
otus <- paste("OTU", sprintf("%05d", seq(1:length(gamma1))), sep = "")

# Initiate Communities
site1 <- sample(otus, size = 100000, replace = T, prob = gamma1)
site2 <- sample(otus, size = 100000, replace = T, prob = gamma2)
site3 <- sample(otus, size = 100000, replace = T, prob = gamma3)

site1.t <- table(site1); site2.t <- table(site2); site3.t <- table(site3)

specnumber(site1.t)
specnumber(site2.t)
specnumber(site3.t)

size.comm <- 1000000
props <- seq(size.comm/100, size.comm, 50000)

ratio.tab <- matrix(NA, nrow = 5, ncol = length(props))
rownames(ratio.tab) <- c("Null", "Uneven", "Even", "Same", "Proportion")

bc.tab <- matrix(NA, nrow = 4, ncol = length(props))
rownames(bc.tab) <- c("1-2", "1-3", "1-4", "Proportion")

for(i in 1:length(props)){
  n <- props[i]
  # Initialize
  temp.null <- c(sample(otus, size = size.comm, replace = T, prob = gamma1))
  temp.uneven <- c(sample(otus, size = size.comm, replace = T, prob = gamma1),
                   sample(otus, size = n, replace = T, prob = gamma2))
  temp.even <- c(sample(otus, size = size.comm, replace = T, prob = gamma1),
                 sample(otus, size = n, replace = T, prob = gamma3))
  temp.same <- c(sample(otus, size = size.comm, replace = T, prob = gamma1),
                 sample(otus, size = n, replace = T, prob = gamma1))
  tab1 <- table(temp.null)
  tab2 <- table(temp.uneven)
  tab3 <- table(temp.even)
  tab4 <- table(temp.same)

  sbys <- matrix(NA, ncol = length(otus), nrow = 4)
  colnames(sbys) <- otus
  rownames(sbys) <- c("site1", "site2", "site3",
                    "site4")

  for(j in 1:dim(sbys)[2]){
    if (otus[j] %in% names(tab1)){
      sbys[1, j] <- tab1[which(names(tab1) == otus[j])]
    } else {sbys[1, j] <- 0}

    if (otus[j] %in% names(tab2)){
      sbys[2, j] <- tab2[which(names(tab2) == otus[j])]
    } else {sbys[2, j] <- 0}

    if (otus[j] %in% names(tab3)){
      sbys[3, j] <- tab3[which(names(tab3) == otus[j])]
    } else {sbys[3, j] <- 0}

    if (otus[j] %in% names(tab4)){
      sbys[4, j] <- tab4[which(names(tab4) == otus[j])]
    } else {sbys[4, j] <- 0}

  }


  sbys2 <- rrarefy(sbys, 10000)

  rich <- specnumber(sbys2)

  ratio.tab[1, i] <- round(rich[1]/rich[1], 2)
  ratio.tab[2, i] <- round(rich[2]/rich[1], 2)
  ratio.tab[3, i] <- round(rich[3]/rich[1], 2)
  ratio.tab[4, i] <- round(rich[4]/rich[1], 2)
  ratio.tab[5, i] <- props[i] / size.comm

  dis <- as.matrix(vegdist(sbys2, method = "bray"))
  bc.tab[1, i] <- dis[2, 1]
  bc.tab[2, i] <- dis[3, 1]
  bc.tab[3, i] <- dis[4, 1]
  bc.tab[4, i] <- props[i] / size.comm

  print(paste(i, " of ", dim(ratio.tab)[2], sep = ""))
}

ratio.tab.even <- ratio.tab
bc.tab.even <- bc.tab

```

### Species Abundance Distrubtions
```{r}
layout(matrix(1:3, ncol = 3))
par(mar = c(4, 2, 4, 1), oma = c(0, 3, 1, 1))
plot(log10(site1.t[rev(order(site1.t))]), col = "blue", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 2, "Log Abundance", line = 3)
mtext(side = 3, "Intact\n", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site2.t[rev(order(site2.t))]), col = "gray40",ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 1, "Species Index (OTU)", line = 1.5)
mtext(side = 3, "Species Abundance Distributions", line = 1.5)
mtext(side = 3, "Less Even\nRelic", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site3.t[rev(order(site3.t))]), col = "gray60", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 3, "More Even\nRelic", line = -2, cex = 0.8)

box(lwd = 1.5, bty = "l")
```

## Different S
```{r}
gamma1 <- rlnorm(n=10000, meanlog = 1, sdlog = 0.98)
gamma1 <- gamma1[rev(order(gamma1))]
gamma2 <- rlnorm(n=50000, meanlog = 1, sdlog = 0.98)
gamma2 <- gamma2[rev(order(gamma2))]
gamma3 <- rlnorm(n=5000, meanlog = 1, sdlog = 0.98)
gamma3 <- gamma3[rev(order(gamma3))]


plot(log10(gamma1), col = "forestgreen",
     main = "Species Abundance Distributions",
     xlab = "Species Index", ylab = "Log Probability",
     las = 1)
points(log10(gamma2), col = "red")
points(log10(gamma3), col = "blue")

# Define OTUs
otus1 <- paste("OTU", sprintf("%05d", seq(1:length(gamma1))), sep = "")
otus2 <- paste("OTU", sprintf("%05d", seq(1:length(gamma2))), sep = "")
otus3 <- paste("OTU", sprintf("%05d", seq(1:length(gamma3))), sep = "")

otus <- otus2

# Initiate Communities
site1 <- sample(otus1, size = 100000, replace = T, prob = gamma1)
site2 <- sample(otus2, size = 100000, replace = T, prob = gamma2)
site3 <- sample(otus3, size = 100000, replace = T, prob = gamma3)

site1.t <- table(site1); site2.t <- table(site2); site3.t <- table(site3)

specnumber(site1.t)
specnumber(site2.t)
specnumber(site3.t)

size.comm <- 1000000
props <- seq(size.comm/100, size.comm, 50000)

ratio.tab <- matrix(NA, nrow = 5, ncol = length(props))
rownames(ratio.tab) <- c("Null", "Larger", "Smaller", "Same", "Proportion")

bc.tab <- matrix(NA, nrow = 4, ncol = length(props))
rownames(bc.tab) <- c("1-2", "1-3", "1-4", "Proportion")

for(i in 1:length(props)){
  n <- props[i]
  # Initialize
  temp.null <- c(sample(otus1, size = size.comm, replace = T, prob = gamma1))
  temp.uneven <- c(sample(otus1, size = size.comm - n, replace = T, prob = gamma1),
                   sample(otus2, size = n, replace = T, prob = gamma2))
  temp.even <- c(sample(otus1, size = size.comm - n, replace = T, prob = gamma1),
                 sample(otus3, size = n, replace = T, prob = gamma3))
  temp.same <- c(sample(otus1, size = size.comm - n, replace = T, prob = gamma1),
                 sample(otus1, size = n, replace = T, prob = gamma1))
  tab1 <- table(temp.null)
  tab2 <- table(temp.uneven)
  tab3 <- table(temp.even)
  tab4 <-table(temp.same)

  sbys <- matrix(NA, ncol = length(otus), nrow = 4)
  colnames(sbys) <- otus
  rownames(sbys) <- c("site1", "site2", "site3",
                    "site4")

  for(j in 1:dim(sbys)[2]){
    if (otus[j] %in% names(tab1)){
      sbys[1, j] <- tab1[which(names(tab1) == otus[j])]
    } else {sbys[1, j] <- 0}

    if (otus[j] %in% names(tab2)){
      sbys[2, j] <- tab2[which(names(tab2) == otus[j])]
    } else {sbys[2, j] <- 0}

    if (otus[j] %in% names(tab3)){
      sbys[3, j] <- tab3[which(names(tab3) == otus[j])]
    } else {sbys[3, j] <- 0}

    if (otus[j] %in% names(tab4)){
      sbys[4, j] <- tab4[which(names(tab4) == otus[j])]
    } else {sbys[4, j] <- 0}

  }

  sbys2 <- rrarefy(sbys, 10000)

  rich <- specnumber(sbys2)

  ratio.tab[1, i] <- round(rich[1]/rich[1], 2)
  ratio.tab[2, i] <- round(rich[2]/rich[1], 2)
  ratio.tab[3, i] <- round(rich[3]/rich[1], 2)
  ratio.tab[4, i] <- round(rich[4]/rich[1], 2)
  ratio.tab[5, i] <- props[i] / size.comm

  dis <- as.matrix(vegdist(sbys2, method = "bray"))
  bc.tab[1, i] <- dis[2, 1]
  bc.tab[2, i] <- dis[3, 1]
  bc.tab[3, i] <- dis[4, 1]
  bc.tab[4, i] <- props[i] / size.comm

  print(paste(i, " of ", dim(ratio.tab)[2], sep = ""))
}

ratio.tab.s <- ratio.tab
bc.tab.s <- bc.tab
```

### Species Abundance Distrubtions
```{r}
ratio.tab <- ratio.tab.s

layout(matrix(1:3, ncol = 3))
par(mar = c(4, 2, 4, 1), oma = c(0, 3, 1, 1))
plot(log10(site1.t[rev(order(site1.t))]), col = "blue", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 2, "Log Abundance", line = 3)
mtext(side = 3, "Intact\n", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site2.t[rev(order(site2.t))]), col = "gray40",ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 1, "Species Index (OTU)", line = 1.5)
mtext(side = 3, "Species Abundance Distributions", line = 1.5)
mtext(side = 3, "Larger S\nRelic", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site3.t[rev(order(site3.t))]), col = "gray60", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 3, "Smaller S\nRelic", line = -2, cex = 0.8)

box(lwd = 1.5, bty = "l")
```

# Manuscript Figures
## Figure 1
```{r}
png(filename="../figures/Figure1-Model.png",
    width = 1600, height = 1300, res = 96*2)

layout(matrix(c(1,1,2,3), 2, 2, byrow=T), heights=c(4, 3))
par(oma = c(0.5, 1, 0.5, 1))

# Conceptual Model
img <- readPNG("../figures/Figure1A.png")
par(mar = c(0, 0, 0, 2))
plot(1:10, type = "n", axes = F)
rasterImage(img, 1, 0.55, 9.65, 10.5)
text(x = 0.8, y = 10.2, "a", cex = 1.5, font = 2)


par(mar = c(4, 4, 1, 5) + 0.5)

# Evenness
ratio.tab <- ratio.tab.even
plot(ratio.tab[5, ], ratio.tab[1, ], type = 'n',
     pch = 23, col = "black", bg = "white", axes = F,
     xlab = "",
     ylab = "",
     xlim = c(0, 1), ylim = c(0.78,1.22), las = 1)

# abline(h=1, lty = 3, lwd = 1.5)

points(ratio.tab[5, ], ratio.tab[2, ], pch = 22,bg = "gray40")
points(ratio.tab[5, ], ratio.tab[3, ], pch = 21, bg = "gray60")
points(ratio.tab[5, ], ratio.tab[4, ], pch = 23,bg = "white")

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1,
     at = c(seq(0.8, 1.4, 0.2)))
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5,
     at = c(seq(0.8, 1.4, 0.2)))
box(lwd = 1.5)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1.25)
mtext(side = 2, "Richness Ratio", line = 2.5, cex = 1.25)

text(x = 1.02, y = 1.18, expression(paste(italic("E")["relic"],
     " > ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 1, expression(paste(italic("E")["relic"],
     " = ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 0.81, expression(paste(italic("E")["relic"],
     " < ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)

text(x = 0.02, y = 1.2, "b", cex = 1.5, font = 2)

# Richness
ratio.tab <- ratio.tab.s
plot(ratio.tab[5, ], ratio.tab[1, ], type = 'n',
     pch = 23, col = "black", bg = "white", axes = F,
     xlab = "",
     ylab = "",
     xlim = c(0, 1), ylim = c(0.77,1.43), las = 1)

# abline(h=1, lty = 3, lwd = 1.5)

points(ratio.tab[5, ], ratio.tab[2, ], pch = 22,bg = "gray40")
points(ratio.tab[5, ], ratio.tab[3, ], pch = 21, bg = "gray60")
points(ratio.tab[5, ], ratio.tab[4, ], pch = 23,bg = "white")

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1,
     at = c(seq(0.8, 1.4, 0.2)))
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5,
     at = c(seq(0.8, 1.4, 0.2)))
box(lwd = 1.5)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1.25)
mtext(side = 2, "Richness Ratio", line = 2.5, cex = 1.25)

text(x = 1.02, y = 1.35, expression(paste(italic("S")["relic"],
     " > ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 1, expression(paste(italic("S")["relic"],
     " = ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 0.85, expression(paste(italic("S")["relic"],
     " < ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)

text(x = 0.02, y = 1.4, "c", cex = 1.5, font = 2)

#legend("topleft", c("Relic has Larger S", "Relic has Smaller S"),
#       pch = c(22, 21), pt.bg = c("gray40", "gray60"), bty = "n",
#       cex = 0.8)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure1-Model.png")
grid.raster(img)

```


## Supplemenatal Figure (SI)
```{r}
png(filename="../figures/FigureS1-SimulationBC.png",
    width = 1800, height = 600, res = 96*2)

layout(matrix(c(1,2), 1, 2, byrow=T))
par(mar = c(4, 4, 1, 5.5) + 0.5, oma = c(0, 1, 1, 1))

# Evenness
bc.tab <- bc.tab.even
plot(bc.tab[4, ], bc.tab[1, ], type = 'n', axes = F,
     xlab = "",
     ylab = "Bray Curtis Dissimilarity",
     xlim = c(0, 1), ylim = c(0.42, 0.51), las = 1)
points(bc.tab[4, ], bc.tab[1, ], pch = 22,bg = "gray40")
points(bc.tab[4, ], bc.tab[2, ], pch = 21, bg = "gray60")
points(bc.tab[4, ], bc.tab[3, ], pch = 23, bg = "white")
abline(h=1, lty = 3, lwd = 1.5)

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5)
box(lwd = 1.5)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1)

text(x = 1.05, y = 0.475, expression(paste("E"["relic"], " < E"["intact"])),
     pos = 4, cex = 1, xpd = T)
text(x = 1.05, y = 0.445, expression(paste("E"["relic"], " = E"["intact"])),
     pos = 4, cex = 1, xpd = T)
text(x = 1.05, y = 0.5, expression(paste("E"["relic"], " > E"["intact"])),
     pos = 4, cex = 1, xpd = T)

#mtext("A", side = 3, outer = T, font = 2, at = 0, padj = 2)
text(x = 0.022, y = 0.505, "a", cex = 1.25, font = 2)


# Richness
bc.tab <- bc.tab.s
plot(bc.tab[4, ], bc.tab[1, ], type = 'n', axes = F,
     xlab = "",
     ylab = "Bray Curtis Dissimilarity",
     xlim = c(0, 1), ylim = c(0.4, 0.58), las = 1)

points(bc.tab[4, ], bc.tab[1, ], pch = 22,bg = "gray40")
points(bc.tab[4, ], bc.tab[2, ], pch = 21, bg = "gray60")
points(bc.tab[4, ], bc.tab[3, ], pch = 23, bg = "white")

abline(h=1, lty = 3, lwd = 1.5)

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5)
box(lwd = 1.5)

text(x = 1.05, y = 0.55, expression(paste("S"["relic"], " > S"["intact"])),
     pos = 4, cex = 1, xpd = T)
text(x = 1.05, y = 0.455, expression(paste("S"["relic"], " = S"["intact"])),
     pos = 4, cex = 1, xpd = T)
text(x = 1.05, y = 0.43, expression(paste("S"["relic"], " < S"["intact"])),
     pos = 4, cex = 1, xpd = T)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1)

#mtext("B", side = 3, outer = T, font = 2, at = 0.5, padj = 2)
text(x = 0.022, y = 0.57, "b", cex = 1.25, font = 2)


# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/FigureS1-SimulationBC.png")
grid.raster(img)
```

```{r}
# Load Regional Pool
gamma1 <- rlnorm(n=10000, meanlog = 1, sdlog = 0.98)
gamma1 <- gamma1[rev(order(gamma1))]
gamma2 <- rlnorm(n=10000, meanlog = 1, sdlog = 2)
gamma2 <- gamma2[rev(order(gamma2))]
gamma3 <- rlnorm(n=10000, meanlog = 1, sdlog = 0.1)
gamma3 <- gamma3[rev(order(gamma3))]

# Simple Plot
layout(1)
plot(log10(gamma1), col = "forestgreen",
     main = "Species Abundance Distributions",
     xlab = "Species Index", ylab = "Log Probability",
     las = 1)
points(log10(gamma2), col = "red")
points(log10(gamma3), col = "blue")

```

# Model With Simulation Output
```{r}
load("../data/simSAD.RData")

# Same Rates
total1 <- table(c(same.rates[[1]], same.rates[[2]]))
active1 <- table(same.rates[[1]])
relic1 <- table(same.rates[[2]])

total1 <- total1[rev(order(total1))]
active1 <- active1[rev(order(active1))]
relic1 <- relic1[rev(order(relic1))]

# Low Decay
total2 <- table(c(low.decay[[1]], low.decay[[2]]))
active2 <- table(low.decay[[1]])
relic2 <- table(low.decay[[2]])

total2 <- total2[rev(order(total2))]
active2 <- active2[rev(order(active2))]
relic2 <- relic2[rev(order(relic2))]

# High Decay
total3 <- table(c(high.decay[[1]], high.decay[[2]]))
active3 <- table(high.decay[[1]])
relic3 <- table(high.decay[[2]])

total3 <- total3[rev(order(total3))]
active3 <- active3[rev(order(active3))]
relic3 <- relic3[rev(order(relic3))]
```

# Model With Simulation Output
```{r}
load("../data/simSAD.RData")

length(same.rates[[1]])

# Same Rates
total1 <- table(sample(c(same.rates[[1]], same.rates[[2]]), 9000))
active1 <- table(sample(same.rates[[1]], 9000))
relic1 <- table(sample(same.rates[[2]], 9000))

total1 <- total1[rev(order(total1))]
active1 <- active1[rev(order(active1))]
relic1 <- relic1[rev(order(relic1))]

# Low Decay
total2 <- table(sample(c(low.decay[[1]], low.decay[[2]]), 9000))
active2 <- table(sample(low.decay[[1]], 9000))
relic2 <- table(sample(low.decay[[2]], 9000))

total2 <- total2[rev(order(total2))]
active2 <- active2[rev(order(active2))]
relic2 <- relic2[rev(order(relic2))]

# High Decay
total3 <- table(sample(c(high.decay[[1]], high.decay[[2]]), 9000))
active3 <- table(sample(high.decay[[1]], 9000))
relic3 <- table(sample(high.decay[[2]], 9000))

total3 <- total3[rev(order(total3))]
active3 <- active3[rev(order(active3))]
relic3 <- relic3[rev(order(relic3))]
```

# Model With Simulation Output
```{r}
load("../data/simSAD.RData")

# Same Rates
total1 <- table(c(same.rates[[1]], same.rates[[2]]))
active1 <- table(same.rates[[1]])
relic1 <- table(same.rates[[2]])

total1 <- total1[rev(order(total1))]
active1 <- active1[rev(order(active1))]
relic1 <- relic1[rev(order(relic1))]

# Low Decay
total2 <- table(c(low.decay[[1]], low.decay[[2]]))
active2 <- table(low.decay[[1]])
relic2 <- table(low.decay[[2]])

total2 <- total2[rev(order(total2))]
active2 <- active2[rev(order(active2))]
relic2 <- relic2[rev(order(relic2))]

# High Decay
total3 <- table(c(high.decay[[1]], high.decay[[2]]))
active3 <- table(high.decay[[1]])
relic3 <- table(high.decay[[2]])

total3 <- total3[rev(order(total3))]
active3 <- active3[rev(order(active3))]
relic3 <- relic3[rev(order(relic3))]
```

# SAD Plots
```{r}
png(filename="../figures/Figure1-SADs.png",
    width = 1600, height = 800, res = 96*2)

# Simple Plot
layout(1)
par(mar = c(3, 6, 2, 2))
plot(log10(active1), col = "gray",
     main = "",
     xlab = "", ylab = "", 
     xlim = c(0, max(length(active1), length(relic2))),
     ylim = c(0, max(max(log10(active1)), max(log10(relic2)), 2.5)),
     las = 1, type = "p", xaxt = "n", yaxt = "n", pch = 16)
points(log10(relic1), col = "gray30", type = "p", pch = 16)
points(log10(relic2), col = "gray30", type = "p", pch = 15)
points(log10(relic3), col = "gray30", type = "p", pch = 16)
legend("topright", legend = c("Active", "Relic Same Rates", 
       "Relic Low Decay", "Relic High Decay"), pch = c(16, 15, 16, 17), 
       bty = "n", col = c("gray", "gray30", "gray30", "gray30"))
axis(side = 1, labels = F, lwd = 2, tck = 0)
axis(side = 2, labels = T, lwd = 2, las = 1, cex.axis = 1.25)
axis(side = 2, labels = F, lwd = 2, tck = 0.02)
mtext(expression(paste("Log"[10], " Abundance")), side = 2, cex = 2, line = 3)
mtext("Species Index", side = 1, cex = 2, line = 1)
box(lwd = 2)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure1-SADs.png")
grid.raster(img)

```

# Simulation with New SAD
```{r}
active.prob <- active1 / sum(active1)
relic1.prob <- relic1 / sum(relic1)    # Same
relic2.prob <- relic2 / sum(relic2)    # Low
relic3.prob <- relic3 / sum(relic3)    # Low

active1.n <- names(active1)
active2.n <- names(active2)
active3.n <- names(active3)
relic1.n <- names(relic1)  # Same
relic2.n <- names(relic2)  # Low 
relic3.n <- names(relic3)  # High


size.comm <- 1000
n.relic <- seq(size.comm/10, size.comm, 50)
n.active <- size.comm - n.relic
props <- n.relic/size.comm

ratio.tab <- matrix(NA, nrow = 4, ncol = length(props))
rownames(ratio.tab) <- c("Same", "Low", "High", "Proportion")

bc.tab <- matrix(NA, nrow = 4, ncol = length(props))
rownames(bc.tab) <- c("1-2", "1-3", "1-4", "Proportion")

for(i in 1:length(props)){

  # Define Community Sizes
  n1 <- n.active[i]
  n2 <- n.relic[i]

  # Simuate Mixing Mdel
  temp.null.same <- c(sample(same.rates[[1]], size = size.comm, replace = F))
  temp.same <- c(sample(same.rates[[1]], size = n1, replace = F),
                 sample(same.rates[[2]], size = n2, replace = F))
  
  temp.null.low <- c(sample(low.decay[[1]], size = size.comm, replace = F))
  temp.low <- c(sample(low.decay[[1]], size = n1, replace =F),
                 sample(low.decay[[2]], size = n2, replace = F))
  
  temp.null.high <- c(sample(high.decay[[1]], size = size.comm, replace = F))
  temp.high <- c(sample(high.decay[[1]], size = n1, replace = F),
                 sample(high.decay[[2]], size = n2, replace = F))
  
  taba <- table(sample(temp.null.same, 1000))
  tabb <- table(sample(temp.null.low, 1000))
  tabc <- table(sample(temp.null.high, 1000))
  tab1 <- table(sample(temp.same, 1000))
  tab2 <- table(sample(temp.low, 1000))
  tab3 <- table(sample(temp.high, 1000))

  ratio.tab[1, i] <- round(length(tab1)/length(taba), 2)
  ratio.tab[2, i] <- round(length(tab2)/length(tabb), 2)
  ratio.tab[3, i] <- round(length(tab3)/length(tabc), 2)
  ratio.tab[4, i] <- props[i]

  # dis <- as.matrix(vegdist(sbys2, method = "bray"))
  # bc.tab[1, i] <- dis[2, 1]
  # bc.tab[2, i] <- dis[3, 1]
  # bc.tab[3, i] <- dis[4, 1]
  # bc.tab[4, i] <- props[i] / size.comm

  print(paste(i, " of ", dim(ratio.tab)[2], sep = ""))
}

ratio.tab.sim <- ratio.tab
bc.tab.sim <- bc.tab

```

### Species Abundance Distrubtions
```{r}
layout(matrix(1:3, ncol = 3))
par(mar = c(4, 2, 4, 1), oma = c(0, 3, 1, 1))
plot(log10(site1.t[rev(order(site1.t))]), col = "blue", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 2, "Log Abundance", line = 3)
mtext(side = 3, "Intact\n", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site2.t[rev(order(site2.t))]), col = "gray40",ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 1, "Species Index (OTU)", line = 1.5)
mtext(side = 3, "Species Abundance Distributions", line = 1.5)
mtext(side = 3, "Less Even\nRelic", line = -2, cex = 0.8)
box(lwd = 1.5, bty = "l")
plot(log10(site3.t[rev(order(site3.t))]), col = "gray60", ylim = c(0, 3.5),
     main = "", axes = F,
     xlab = "", ylab = "",
     las = 1, type = "p", pch = 21, cex = 0.5)
axis(side = 1, labels = F, tick = F)
axis(side = 2, labels = T, las = 1)
mtext(side = 3, "More Even\nRelic", line = -2, cex = 0.8)

box(lwd = 1.5, bty = "l")
```

# Manuscript Figures
## Figure 1
```{r}
png(filename="../figures/Figure1-Model_NEW.png",
    width = 1600, height = 1300, res = 96*2)

layout(matrix(c(1,1,2,3), 2, 2, byrow=T), heights=c(4, 3))
par(oma = c(0.5, 1, 0.5, 1))

# Conceptual Model
img <- readPNG("../figures/Figure1A.png")
par(mar = c(0, 0, 0, 2))
plot(1:10, type = "n", axes = F)
rasterImage(img, 1, 0.55, 9.65, 10.5)
text(x = 0.8, y = 10.2, "a", cex = 1.5, font = 2)


par(mar = c(4, 4, 1, 5) + 0.5)

# Evenness
ratio.tab <- ratio.tab.sim
plot(ratio.tab[5, ], ratio.tab[1, ], type = 'n',
     pch = 23, col = "black", bg = "white", axes = F,
     xlab = "",
     ylab = "",
     xlim = c(0, 1), ylim = c(0.78,1.22), las = 1)

# abline(h=1, lty = 3, lwd = 1.5)

points(ratio.tab[5, ], ratio.tab[2, ], pch = 22,bg = "gray40")
points(ratio.tab[5, ], ratio.tab[3, ], pch = 21, bg = "gray60")
points(ratio.tab[5, ], ratio.tab[4, ], pch = 23,bg = "white")

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1,
     at = c(seq(0.8, 1.4, 0.2)))
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5,
     at = c(seq(0.8, 1.4, 0.2)))
box(lwd = 1.5)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1.25)
mtext(side = 2, "Richness Ratio", line = 2.5, cex = 1.25)

text(x = 1.02, y = 1.18, expression(paste(italic("E")["relic"],
     " > ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 1, expression(paste(italic("E")["relic"],
     " = ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 0.81, expression(paste(italic("E")["relic"],
     " < ", italic("E")["intact"])), pos = 4, cex = 1, xpd = T)

text(x = 0.02, y = 1.2, "b", cex = 1.5, font = 2)

# Richness
ratio.tab <- ratio.tab.s
plot(ratio.tab[5, ], ratio.tab[1, ], type = 'n',
     pch = 23, col = "black", bg = "white", axes = F,
     xlab = "",
     ylab = "",
     xlim = c(0, 1), ylim = c(0.77,1.43), las = 1)

# abline(h=1, lty = 3, lwd = 1.5)

points(ratio.tab[5, ], ratio.tab[2, ], pch = 22,bg = "gray40")
points(ratio.tab[5, ], ratio.tab[3, ], pch = 21, bg = "gray60")
points(ratio.tab[5, ], ratio.tab[4, ], pch = 23,bg = "white")

axis(side = 1, labels = T, tck = -0.05, lwd = 1.5, las = 1)
axis(side = 2, labels = T, tck = -0.05, lwd = 1.5, las = 1,
     at = c(seq(0.8, 1.4, 0.2)))
axis(side = 1, labels = F, tck = 0.02, lwd = 1.5)
axis(side = 2, labels = F, tck = 0.02, lwd = 1.5,
     at = c(seq(0.8, 1.4, 0.2)))
box(lwd = 1.5)

mtext(side = 1, "Proportion of Relic DNA", line = 2.5, cex = 1.25)
mtext(side = 2, "Richness Ratio", line = 2.5, cex = 1.25)

text(x = 1.02, y = 1.35, expression(paste(italic("S")["relic"],
     " > ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 1, expression(paste(italic("S")["relic"],
     " = ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)
text(x = 1.02, y = 0.85, expression(paste(italic("S")["relic"],
     " < ", italic("S")["intact"])), pos = 4, cex = 1, xpd = T)

text(x = 0.02, y = 1.4, "c", cex = 1.5, font = 2)

#legend("topleft", c("Relic has Larger S", "Relic has Smaller S"),
#       pch = c(22, 21), pt.bg = c("gray40", "gray60"), bty = "n",
#       cex = 0.8)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure1-Model_NEW.png")
grid.raster(img)

```




